---
title: "Correlation between Forearm Length and Height"
author: "CS5302 (.)"
date: "September 26, 2020 (.)"
output: pdf_document
---

## I. Introduction 

Part 1 (.)

Following statistical inference was conducted to learn about how a person's 
forearm length is related to his/her height. Unit for both variables are fixed 
as *centimeters (cm)*. \

The sample data was collected from nine STA302 students in a randomly assigned 
breakout group (group 5). Samples variate in gender 
and ethnicity.\

Here, **explanatory variable** is *forearm length*, and **response variable** 
is *height*. Our goal explains the reason of variable choice: to predict a 
person's height based on his/her forearm length. In real life, this model
could be used to predict a person's height if they can only measure a forearm 
length (short ruler).


## II. Exploratory Data Analysis

Part 2 (.)

### Explanatory variable distribution 
```{r, include = FALSE, out.width="50%"}
library(tidyverse)
```

```{r, echo = FALSE, out.width="50%", fig.align='center'}
height <- c(167, 180, 170, 183, 180, 155, 176, 168, 160)
forearm <- c(24, 30, 26, 30, 30, 23, 28, 24, 23)
f_to_h <- data.frame(forearm, height)

ggplot(f_to_h, aes(x=forearm)) + ggtitle("5302 Histogram of Forearm Length (cm)")+
  geom_histogram(binwidth = 1, fill = "darkorchid4")

```
```{r, include = FALSE}
summary(f_to_h$forearm)
```

The explanatory variable (forearm length) is not at all normally distributed. 
The distribution rather looks like the opposite of normal with two modes at
each tail. Although the mean of 26.44 cm and the median of 26 cm tell us that 
the distribution is quite centered, it is clear from the plot that it has two
modes at two opposite ends. There is no observable outlier. 


### Response variable distribution

```{r, echo = FALSE, out.width="50%", fig.align='center'}
ggplot(f_to_h, aes(x=height)) + ggtitle("5302 Histogram of Height (cm)")+
  geom_histogram(binwidth = 2, fill = "khaki4")
```
```{r, include = FALSE}
summary(f_to_h$height)
```


The response variable (height) is not normally distributed. Instead, with mean
of 171 cm, median of 170 cm, and mode of 180 cm (with 2 data points), it is 
closer to a uniform distribution than a normal distribution. It ranges from 155 
cm to 183 cm with no noticeable outlier or skewness. 


### Both explanatory and response variables

```{r, echo = FALSE, out.width="70%", fig.align='center'}
sp <- ggplot(f_to_h, aes(x=forearm, y=height)) + 
  ggtitle("5302 Scatterplot of Forearm length vs. Height ") + geom_point(shape = 15)
sp + geom_point() 
```

From the scatterplot above, it seems like there exists a positive linear 
relationship between forearm length and height. To be more precise, in the next 
part, we will conduct hypothesis testing and regression analysis on our sample 
data. 


## III. Methods and Model

Part 3 (.)

```{r, echo = FALSE,fig.align='center',out.width="70%"}
sp2<- sp+ geom_point() + 
  stat_smooth(method=lm, se = FALSE, colour = "navajowhite3")
sp2
```


### Hypotheses 
**$H_0$** = There is no linear relationship between forearm length and height. 
($\beta_1=0$) \
**$H_1$** = There is some linear relationship between forearm length and
height. ($\beta_1\neq 0$)

```{r, include = FALSE}
lfit = lm(f_to_h$height ~ f_to_h$forearm)
summary(lfit)
```


### Results

After conducting linear regression analysis on the sample data, these are some
of our numerical outputs: \
\begin{center}
\begin{tabular}{ |c|c| } 
 \hline
 Intercept($b_0$) & 92.94  \\ 
 \hline
 Slope($b_1$) & 2.95  \\
 \hline
 p-value of $b_0$ & 2.69e-05 \\
 \hline
 p-value of $b_1$ & 8.03e-05 \\
 \hline
\end{tabular}
\end{center}

Hence the **fitted line** is: 
$$
\hat{y} = 92.94 + 2.95x
$$
where $\hat{y}$ is the predicted height (cm) given x, forearm length (cm).\ 
\

### Conclusion

With our significance level of 5%, since the p-value of our slope parameter 
is very small and $b_1=2.95>0$, there is a 
*very strong evidence against the null hypothesis*. \
Thus, we conclude that it is very likely that the population parameter
is not zero ($\beta_1\neq0$), meaning that there is a positive relationship 
between a person's forearm length and height. \
\
The interpretation of $b_1$ is the change in height in cm when the 
forearm length increases by 1 cm. That means for every 1cm increase in 
forearm length, there is 2.95 cm of positive change in height.\
The interpretation of $b_0$ is the estimated height in cm when the 
forearm length is zero. In a mathematical context, $b_0=92.94$ indicates that 
when one's forearm length is 0cm the height is predicted to be 92.94cm, but 
in the context of human body, when one's forearm length is 0cm, then the height 
should also be 0, or at least very close to 0. 


## IV. Discussions and Limitations

Part 4 (.)

### Potential Lurking Variables and limitations
Special types of physical activities could be a possible lurking variable. 
Certain physical activities could help a human body develop with a slightly 
different proportion of body parts, such as longer arms or longer legs. Also, 
genetics could contribute on the shape and proportion of a body. \
\
It is hard to say that our sample data was unbiased.\
First, the *race and gender distribution was unclear, and samples' age were not diverse*. 
In fact, several 
physiological studies has shown ethnic/ gender differences in body proportions. 
When a person gets older, his/her height shrinks due to the transform of backbone, 
but the forearm length changes less. 
But until there is a strict cause and effect relationship proven, these remain 
as a possible hypothesis.\
In addition, *our number of sample data was not large enough* that the 
randomness has big impact on our model. In the future, we could increase the
size of data and fit a more reliable model. 

### Another pair of variables

In the future, it would be interesting to conduct a linear regression 
analysis on the relationship between the number of university courses taking 
at the same time versus one's happiness. \
In this case, the number of university courses taking at the same time
would be an independent variable, and the happiness index would be a response 
variable. 
