---
title: "MLR Analysis on GTA House Price (.)"
author: "Chaerin Song, Id 1005745302 (.)"
date: "December 5, 2020 (.)"
output:
  pdf_document: default
  word_document: default
---

## I. Data Wrangling

Part 1 (.)\

*House price data was obtained from the*
*Toronto Real Estate Board (TREB) on detached houses in the city of Toronto *
*and the city of Mississauga.*

```{r, include= FALSE}
library(broom)
library(tidyverse)
library(cowplot)

# Randomizing data
set.seed(1005745302)

# Reading the data
csv5302 <- read.csv("real203.csv")
house_data5302 <- sample_n(csv5302, size = 150)
```

### Selecting random 150 cases
These are the sorted ID's of selected 150 samples.
```{r, echo=FALSE}
# Sorting and selecting only the ID's to display
ID5302 <- house_data5302[,1]
house_data5302<- house_data5302[order(house_data5302$ID),]
house_data5302$ID
```

### Creating a new variable
And we hereby transformed "lotwidth" and "lotlength" predictors into a new predictor "lotsize."
Lotsize will be representing the size of the house in square feet.
```{r, echo=FALSE}
# create a new variable called lotsize by multiplying lotlnegth and lotwidth.
house_data5302 <- house_data5302 %>% 
  mutate(lotsize = lotwidth + lotlength)
house_data5302 <- subset(house_data5302, select = -c(lotlength, lotwidth))
head(house_data5302)
```

### Removing a predictor and (up to 11) cases

We will remove "maxsqfoot" predictor from our data since it has too many NA values. 

```{r, echo = FALSE}
house_data5302 <- subset(house_data5302, select = -c(maxsqfoot))
head(house_data5302, 3)
```


Now, we will remove every case that has at least one N/A value, 
which counts to be 10 observations.
```{r, echo = FALSE}
print("Number of NA rows:")
sum(!complete.cases(house_data5302))
house_data5302 <- na.omit(house_data5302)
```



We can calculate leverage and Cook's distance for each point with a full numerical 
additive model. \

In fact, there are many points that exceed the threshold and are categorized as 
leverage points, while there is no point with extremely high Cook's distance. 
However, we can see that the case of index 106 has a noticeably higher Cook's distance 
than other points; in addition, it also has the biggest hat value.\
We could remove this case from our data.
```{r, echo = FALSE, message = FALSE, warning = FALSE}
numeric_data <- subset(house_data5302, select = -c(ID, location))
attach(numeric_data)
m_demo5302 <- lm(sale~list+bedroom+bathroom+parking+taxes+lotsize)
pp1=length(coef(m_demo5302))
threshold=2*(pp1)/length(sale)
hii=hatvalues(m_demo5302)
cooks<-cooks.distance(m_demo5302)

print("Top 5 Cooks distance:")
head(round(sort(cooks, decreasing=TRUE)[1:12], 3),5)
print("Top 5 hat values that exceeds the threshold:")
head(round(sort(hii, decreasing=TRUE)[1:12], 3),5)

detach(numeric_data)
```

```{r, include = FALSE}
house_data5302 <- house_data5302[-c(106), ]
```


## II. Exploratory Data Analysis
Part 2 (.)\
Table (.)

### Variable Classification 

**Type of Variables**
\begin{center}
\begin{tabular}{ |c|c| } 
 \hline
  Variable & Type   \\ 
 \hline
 \hline
 sale & continuous \\
 
 list & continuous \\ 
 
 bedroom & discrete \\

 bathroom & discrete \\
 
 parking & discrete \\
 
 taxes & continuous \\
 
 lotsize & continuous \\
 
 location & categorical \\
 \hline
\end{tabular}
\end{center}

### Pairwise Correlation and Scatterplot Matrix

The following matrix plot demonstrates a pairwise correlation and scatterplot of two variables.
Variables that are included are: *sale (response variable), list, bedroom, bathroom, parking, * *taxes, and lotsize. *

The upper diagonal half of the panel demonstrates scatterplots of two corresponding variables, 
and the lower half demonstrates pairwise correlation of these variables. 


```{r, echo = FALSE}

numeric_data <- subset(house_data5302, select = -c(ID, location))

# Correlation panel
panel.cor <- function(x, y){
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- round(cor(x, y, use = "complete.obs"), digits=2)
    txt <- paste0("R = ", r)
    cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt)
}

# Customize upper panel
upper.panel<-function(x, y){
  points(x,y, pch = 19)
}
# Create the plots
pairs(numeric_data,
      lower.panel = panel.cor,
      upper.panel = upper.panel,
      main = "Pairwise Correlation and Scatterplot Matrix 5302")

```
Based on the matrix above, we can rank each quantitative predictor based on their correlation coefficient with the sale variable. \
Here is the summary:

**Each Quantitative Predictor for Sale Price Rank (base on R value)**
\begin{center}
\begin{tabular}{ |c c|c| } 
 \hline
  Variable & Type & Cor.coeff  \\ 
 \hline
 \hline
 1 & list & R = 0.99 \\ 
 
 2 & taxes & R = 0.71 \\
 
 3 & bathroom & R = 0.63 \\
  
 4 & bedroom & R = 0.47 \\

 5 & lotsize & R = 0.31 \\
 
 6 & parking & R = 0.08 \\

 \hline
\end{tabular}
\end{center}

Let's see if there is any multicollinearity between any two predictors. 
Although there are some predictors that have a visibly positive linear relationship, 
none of their pairwise correlation value indicates that they are strongly correlated. \

Based on the scatterplot matrix, it seems like sale~taxes violates the homoscedasticity assumption the most. We shall check this by looking at the scale-location plot. 
```{r, out.width="70%", message=FALSE, echo = FALSE, fig.align='center'}

ST5302 = lm(house_data5302$sale ~ house_data5302$taxes)
plot(ST5302, 3, main = "Scale-Location plot 5302")
```
Since the graph is far from being horizontal, we can confirm that taxes as a predictor of sale price strongly violates the assumption of constant variance. 

## III. Methods and Model

Part 3 (.)

Table (.)

### Fitting a full model

Before we go ahead and fit an additive linear regression model, we shall 
assign numerical values for location predictor, so that it will be an 
indicator variable; 
1 for T(Toronto), 0 for (Mississauga).
```{r, echo = FALSE}
# house_data5302[house_data5302$location == "T"] <- 1
house_data5302$location <- recode(house_data5302$location, T = 1, M = 0)
head(house_data5302)
```

Now we will fit an additive linear regression model for sale price, including our 
recoded location variable.\ 
Here is the summary of the model:

```{r, message = FALSE, echo = FALSE}
attach(house_data5302)
m1_5302 <- lm(sale~list+bedroom+bathroom+parking+taxes+lotsize+location)
summary(m1_5302)
```

Our fitted model is: \
$\hat{y} = 79020 + 0.8333x_1 + 19900x_2 - 612.0x_3 -12030x_4 + 20.97x_5 -73.62x_6 + 82950 x_7$

where each explanatory variable represents: \
\
$x_1$ = list price in CAD\
$x_2$ = number of bedrooms\
$x_3$ = number of bathrooms\
$x_4$ = number of parking spots\
$x_5$ = taxes in CAD\
$x_6$ = lot size in square feet\
$x_7$ = 1 for Toronto neighborhood, and 0 for Missisauga neighborhood.\
\

Our global F-test turned out to be significant with an extremely small p-value.\
Now, let's look at the estimated regression coefficients and the 
p-values for individual t-tests.


**Full additive fitted model**
\begin{center}
\begin{tabular}{ |c|c c| } 
 \hline
  Predictor & Est. Coeff. & p-value  \\ 
 \hline
 \hline
 list & 0.8333 & 0.0000 \\ 
 
 bedroom & 19900. & 0.1924 \\
 
 bathroom & - 612.0 & 0.9641 \\
  
 parking & -12030 & 0.1452 \\

 taxes & 20.97 & 0.0000 \\
 
 lotsize & -73.62 & 0.7838 \\
 
 location & 82950. & 0.0285 \\

 \hline
\end{tabular}
\end{center}


With a benchmark significance level of 5%, list price, taxes, and location can 
be used as factors to help predict the sale price 
*over and above* other predictors. \
Because our model's global F-test is significant and some of the t-tests are 
significant, we can assume that there are some useful explanatory variables in 
our model for predicting the sale price. 


### Stepwise regression with AIC

Now, we are going to use a stepwise regression with backward AIC from our 
original model. 

```{r, include = FALSE}
AIC_5302 <- step(m1_5302, direction = "backward")
summary(AIC_5302)
```


Our final model is: \
$\hat{y} = 71150 + 0.8296x_1 + 19760x_2 - 12930x_3 + 20.68x_4 + 85090x_5$\
where each explanatory variable represents: \
\
$x_1$ = list price in CAD\
$x_2$ = number of bedrooms\
$x_3$ = number of parking spots\
$x_4$ = taxes in CAD\
$x_5$ = 1 for Toronto neighborhood, and 0 for Mississauga neighborhood.\


In this model, there are only 
5 predictors used to predict the sale price: bedroom, parking, location, taxes, and list.\

Below is the table demonstrating estimated regression coefficients and p-values 
for these four predictors.

**Fitted model using backward AIC**
\begin{center}
\begin{tabular}{ |c|c c| } 
 \hline
  Predictor & Est. Coeff. & p-value \\ 
 \hline
 \hline
 list & 0.8296 & 0.0000 \\ 
 
 bedroom & 19760. & 0.1584 \\
 
 parking & - 12930. & 0.0846 \\

 taxes & 20.68 & 0.0000 \\

 location & 85090. & 0.0156 \\

 \hline
\end{tabular}
\end{center}

This model is not only different in the number of predictors from our original model, 
but also in its common predictors’ estimated coefficients. \
In short, the results are not consistent with our original (full) model. 



### Stepwise regression with BIC

Now, we will use BIC to repeat our what we did above. 

```{r, include = FALSE}
BIC_5302<- step(m1_5302, direction = "backward", k=log(141))
summary(BIC_5302)
```

Our final model is: \
$\hat{y} = 76440 + 0.8313x_1 + 21.21x_2 + 125000x_3$\
where each explanatory variable represents: \
\
$x_1$ = list price in CAD\
$x_2$ = taxes in CAD\
$x_3$ = 1 for Toronto neighborhood, and 0 for Mississauga neighborhood.\
\

In this model, there are only 3 predictors used to predict the sale price: 
location, taxes, and list.\

Below is the table demonstrating estimated regression coefficients and p-values 
for these four predictors.

**Fitted model using backward BIC**
\begin{center}
\begin{tabular}{ |c|c c| } 
 \hline
  Predictor & Est. Coeff. & p-value \\ 
 \hline
 \hline
 list & 0.8313 & 0.0000 \\ 
 
 taxes & 21.21 & 0.0000 \\

 location & 125000. & 0.0000 \\

 \hline
\end{tabular}
\end{center}

The results are not consistent with both our original (full) model and AIC model. 
This model has the least predictors and the smallest individual t-test p-values. 
Predictors' estimated coefficients are different from the previous two models 
as well, although with AIC model they aren't numerically too different. \


## IV. Discussions and Limitations

Part 4 (.)

Plot (.)

### Check MLR assumptions

Below are the 4 diagnostic plots for our BIC model. 
```{r, echo = FALSE, fig.align= 'center'}
par(mfrow = c(2, 2))
plot(BIC_5302)
mtext("4 Diagnostic Plots 5302", side = 3, line = -1.1, outer = TRUE, cex = 1.2,
      col = "darkorchid3")
```

1. **Residuals vs Fitted**: We obtained a horizontal line without distinct 
patterns, which is an indication for a linear relationship.\

2. **Normal Q-Q**: Most of the residuals are normally distributed, as 
they follow the straight dashed line. 
However, some points with large absolute theoretical quantiles are far from 
the dashed line, which indicates that the normal error MLR assumption is not 
completely satisfied. \

3. **Scale-Location**: Although there is a positive slope with the 
smaller values of sale price and the points are not too equally spread, 
there is a generally horizontal line. 
We are mostly satisfying the homoscedasticity MLR assumption. \

4. **Residuals vs Leverage**: There are some points with large absolute 
standardized residuals, but none of the points' Cook's distance exceeds 1, 
nor are any of them beyond the dashed lines. \


### Towards our final model

First, as shown in the scatterplot matrix in part (i), some of our explanatory 
variables are not normally distributed. In the future, we could conduct a 
Box-Cox transformation to transform these variables to be close to normally 
distributed. \

Right now, how much have we come to fit a valid model?\
We drew scatterplots of the data and removed a bad influential point. 
We are mostly satisfying our MLR assumptions, and we fit 3 different models to 
end up with a model with 3 predictors, all statistically significant.\
To better satisfy the homoscedasticity assumption, we could possibly use 
the bootstrap for inference and refit the model. \



